Sweat electrolytes and CFTR activity in the era of CFTR modulators

### 2024-10-09 Edith Zemanick/Sweat Test/sweat_test_analysis.qmd
sweat analysis update

```{r setup}
#| include: false
library(tidyverse)
library(readxl)
library(gtsummary)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Test",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Test",
  "Linux" = "/home/tim/OneDrive/Vigers/CF/Edith Zemanick/Sweat Test"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

  df[, c("Genotype1Function", "Genotype2Function")],
  na.rm = T

@@

@@

# Excluding those with no genotype function estimate

```{r}
df %>%
  filter(!is.na(TotalGenotypeFunction)) %>%
  select(
    Diagnosis, Genotypes1, Genotypes2, Genotype1Function, Genotype2Function,
    TotalGenotypeFunction, Na, K, Cl
  ) %>%
  droplevels(.) %>%
  tbl_summary(by = Diagnosis)
```

```{r}
df %>%
  filter(!is.na(TotalGenotypeFunction)) %>%
  ggplot(
    aes(x = Cl, y = TotalGenotypeFunction, color = Diagnosis, group = Diagnosis)
  ) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())

df %>%
  filter(!is.na(TotalGenotypeFunction)) %>%
  ggplot(
    aes(x = Cl, y = TotalGenotypeFunction, color = Diagnosis, group = Diagnosis)
  ) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  ylim(c(0, 20)) +
  theme(axis.ticks.x = element_blank(), axis.text.x = element_blank())
```

>>>>

```{r setup}
#| include: false
library(tidyverse)
library(readxl)
library(gtsummary)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Edith Zemanick/Sweat Test",
  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Edith Zemanick/Sweat Test",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Edith Zemanick/Sweat Test"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

  df[, c("Genotype1Function", "Genotype2Function")]

@@

# Limit to complete data
    PatientID, Diagnosis, Sex, Age.at.Test.in.Years, Na:Cl,
    Genotypes1, Genotype1Function, Genotypes2, Genotype2Function,
    TotalGenotypeFunction
# Set non-CF to 100% CFTR function (200 because 100 for each allele)
df$TotalGenotypeFunction[df$Diagnosis == "Non-CF"] <- 200

@@

```{r}
#| message: false
#| warning: false
df %>%
  filter(Diagnosis == "CF") %>%
  ggplot(
    aes(x = Na, y = TotalGenotypeFunction, color = Diagnosis, group = Diagnosis)
  ) +
  geom_point() +
  theme()
df %>%
  filter(Diagnosis == "CF") %>%
  ggplot(
    aes(x = Cl, y = TotalGenotypeFunction, color = Diagnosis, group = Diagnosis)
  ) +
  geom_point() +
  theme()
df %>%
  filter(Diagnosis == "CF") %>%
  ggplot(
    aes(x = Na / Cl, y = TotalGenotypeFunction, color = Diagnosis, group = Diagnosis)
  ) +
  geom_point() +
  theme()

df %>%
  filter(Diagnosis == "CF") %>%
  ggplot(
    aes(
      x = Na, y = TotalGenotypeFunction, color = Age.at.Test.in.Years,
      group = Diagnosis
    )
  ) +
  geom_point() +
  theme()
df %>%
  filter(Diagnosis == "CF") %>%
  ggplot(
    aes(
      x = Cl, y = TotalGenotypeFunction, color = Age.at.Test.in.Years,
      group = Diagnosis
    )
  ) +
  geom_point() +
  theme()
df %>%
  filter(Diagnosis == "CF") %>%
  ggplot(
    aes(
      x = Na / Cl, y = TotalGenotypeFunction, color = Age.at.Test.in.Years,
      group = Diagnosis
    )
  ) +
  geom_point() +
  theme()
```

```{r}
df %>%
  filter(Diagnosis == "CF") %>%
  mutate(
    `Age group` = cut(Age.at.Test.in.Years, c(-Inf, 18, Inf), right = F),
    `Cl < 60` = Cl < 60
  ) %>%
  select(Na, Cl, TotalGenotypeFunction, `Cl < 60`, `Age group`) %>%
  tbl_summary(by = `Age group`, missing_text = "Missing")
write.csv(df, file = "~/sweat_analysis_data.csv", na = "", row.names = F)
```

Add CRMS change in sweat and change in ratio

Add CF carrier genotypes to this report

Plot just the CF genotype function
  - Are the chloride values matching up with the CFTR function as we'd expect?
